---
title: "Portfolio 1"
---

> The purpose of this is to replicate the data cleaning in SPSS that I do to get this data ready for analysis. The goal will be to (1) get the datafile, (2) remove the survey responses not collected through Prolific, and (3) look for and remove any outliers or other responses we want to exclude from analysis.

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidyr)


```


This is the datafile that I downloaded from Qualtrics.

```{r getdata}

ideal_society <- read.csv("C:\\Users\\Ben\\OneDrive\\Documents\\01Wake Forest\\2025SpringDataScience\\IdealSocietyData\\RRRPILOT_IdealSocietiesConservative Values_February 10, 2025_07.27.csv")

```

I want to filter out responses based on the StartDate variable because any responses entered before Jan 28 2025 were test runs of the suvey and shouuld not be included in the analysis.

Apparently I need to change the format of the StartDate variable to POSIX. 

```{r cleaningdata_1}
ideal_society$StartDate <- as.POSIXct(ideal_society$StartDate, format="%Y-%m-%d %H:%M:%S", tz="UTC")

```


Here I am filtering out responses that were not collected through Prolific.These were tests of the survey that Qualtrics includes in the final data export.

``` {r cleaningdata_2}

ideal_society_real_data <- ideal_society %>%
  filter(as.Date(StartDate)=="2025-01-28") %>%
  filter(Progress != 14)
  
```

I also removed one response that was incomplete, this is the score of "14" under the Progress variable.

Below I removed one response from the data because their free response was not thoughtful enough. They just wrote "5" for our question asking what they would like to see happen in Washington. We still paid them but want to remove them from analysis.

```{r cleaningdata_3}
ideal_society_clean <- ideal_society_real_data %>%
  filter(Q127 != 5)
```


Here's a plot showing a distribution of participants survey response duration. Practicing visualizing data!

```{r plot_duration}

ideal_society_clean <- ideal_society_clean %>%
  mutate(Duration..in.seconds. = as.numeric(Duration..in.seconds.))

ggplot(ideal_society_clean, aes(x = Duration..in.seconds.)) +
  geom_histogram(binwidth = 25, fill = "blue", color = "black") +
  labs(title = "Distribution of Duration", x = "Duration (in seconds)", y = "Frequency") +
  theme_minimal()



```

Checking for any outliers. Per Prolific's data check standards: we don't have to pay anyone who is three SD below the mean. This

```{r outliers}

z_score_duration <- scale(ideal_society_clean$Duration..in.seconds.)
ideal_society_clean$outlier <- abs(z_score_duration) > 3

```

This checked for outliers based on duration it took to complete the survey and while there were three participants where their duration was farther than 3 standard deviations from the mean, they were all greater than the mean and we are more concerned with people speeding through the survey than taking a longer time. 

I was able to replicate my data cleaning through SPSS using R.
