---
title: "Portfolio 1"
---

> The projects should be numbered consecutively (i.e., in the order in which you began them), and should include for each project a description of the goal, the product (computer program, hand graph, computer graph, etc.), the data, and some interpretation. Reports must be reproducible and of high quality in terms of writing, grammar, presentation, etc.

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidyr)

```



```{r getdata}

ideal_society <- read.csv("C:\\Users\\Ben\\OneDrive\\Documents\\RRRPILOT_IdealSocietiesConservative Values_February 10, 2025_07.27.csv")

```


Changing the format of the StartDate variable to POSIX because apparently I need to. 
```{r cleaningdata_1}
ideal_society$StartDate <- as.POSIXct(ideal_society$StartDate, format="%Y-%m-%d %H:%M:%S", tz="UTC")

```


Filtering out responses that were not collected through Prolific.These were tests of the survey that qualtrics includes in the final data export.
``` {r cleaningdata_2}

ideal_society_real_data <- ideal_society %>%
  filter(as.Date(StartDate)=="2025-01-28") %>%
  filter(Progress != 14)
  
```

Removing one response from the data because their free response was not thoughtful enough. They just wrote "5" for our question asking what they would like to see happen in washington. Not cool.

```{r cleaningdata_3}
ideal_society_clean <- ideal_society_real_data %>%
  filter(Q127 != 5)
```


Here's a plot showing a distribution of participants survey response durations. Practicing visualizing data!

```{r plot_duration}

ideal_society_clean <- ideal_society_clean %>%
  mutate(Duration..in.seconds. = as.numeric(Duration..in.seconds.))

ggplot(ideal_society_clean, aes(x = Duration..in.seconds.)) +
  geom_histogram(binwidth = 25, fill = "blue", color = "black") +
  labs(title = "Distribution of Duration", x = "Duration (in seconds)", y = "Frequency") +
  theme_minimal()



```
```{r violinplot}

ggplot(ideal_society_clean, aes(x = "", y = Duration..in.seconds.)) +
  geom_violin(fill = "blue", color = "black") +
  geom_boxplot(width = 0.2, color = "black", outlier.shape = 4)+
  labs(title = "Distribution of Duration", x = "", y = "Duration (in seconds)") +
  theme_minimal()


```
Checking for any outliers. According to Prolific, we don't have to pay anyone who is three SD below the mean.

```{r outliers}

z_score_duration <- scale(ideal_society_clean$Duration..in.seconds.)
ideal_society_clean$outlier <- abs(z_score_duration) > 3

```

This checked for outliers based on duration it took to complete the survey and while there were three participants where their duration was farther than 3 standard deviations from the mean, they were all greater than the mean and we are more concerned with people speeding through the suvey than taking a longer time. 

This cleaned dataset is ready to be used in my analysis!! WHOOOOO!