---
title: "Portfolio 3"
---

The goal of this project is to run a regression analysis on pilot data i collected for my thesis project. I think i will break this up into several portfolios eventually but i need this all here for now.


```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidyr)
library(dplyr)

```

```{r cleaneddata}

ideal_society <- read.csv("C:\\Users\\Ben\\OneDrive\\Documents\\01Wake Forest\\2025SpringDataScience\\IdealSocietyData\\RRRPILOT_IdealSocietiesConservative Values_February 10, 2025_07.27.csv")

ideal_society$StartDate <- as.POSIXct(ideal_society$StartDate, format="%Y-%m-%d %H:%M:%S", tz="UTC")


ideal_society_real_data <- ideal_society %>%
  filter(as.Date(StartDate)=="2025-01-28") %>%
  filter(Progress != 14)

ideal_society_clean <- ideal_society_real_data %>%
  filter(Q127 != 5)

politicalorientation_data <- read.csv("C:\\Users\\Ben\\OneDrive\\Documents\\01Wake Forest\\2025SpringDataScience\\IdealSocietyData\\CopyMergefile.csv")

ideal_society_clean <- ideal_society_clean %>%
  left_join(politicalorientation_data, by = "PROLIFIC_PID")
  
ideal_society_clean <- ideal_society_clean %>%
  rename (poli_ori = PROLIFIC_Political.spectrum..us.)

ideal_society_clean <- ideal_society_clean %>%
  rename(LILH = Q7) %>%
  rename(LIHH = Q4) %>%
  rename(HILH = Q6) %>%
  rename(HIHH = Q5) %>%
  rename(LFLH = Q8) %>%
  rename(LFHH = Q124) %>%
  rename(HFLH = Q125) %>%
  rename(HFHH = Q126) %>%
  rename(LRLH = Q137) %>%
  rename(LRHH = Q139) %>%
  rename(HRLH = Q140) %>%
  rename(HRHH = Q141)

```

Next i want to remove variables I don't need for my analysis. That includes everything except for the sociaty ratings ("LFLH" and so on) and the political ideology measures


```{r removingvariables}
ideal_society_simplified <- ideal_society_clean %>%
  select(LFLH,
         LFHH,
         HFLH,
         HFHH,
         LRLH,
         LRHH,
         HRLH,
         HRHH,
         Q2,
         Q26,
         Q27,
         Q28,
         poli_ori)
```

Now that it is just the simmplified data, i want to zscore the political oreintation measures.
