---
title: "Portfolio 3"
---

The goal of this project is to run a regression analysis on pilot data i collected for my thesis project. 

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidyr)
library(dplyr)
library(lme4)
library(MuMIn)
library(lmerTest)

```

Same process i did from portfolio 1 and 2.

```{r getting-and-cleaning-data}

ideal_society <- read.csv("C:\\Users\\Ben\\OneDrive\\Documents\\01Wake Forest\\2025SpringDataScience\\IdealSocietyData\\RRRPILOT_IdealSocietiesConservative Values_February 10, 2025_07.27.csv")

ideal_society$StartDate <- as.POSIXct(ideal_society$StartDate, format="%Y-%m-%d %H:%M:%S", tz="UTC")


ideal_society_real_data <- ideal_society %>%
  filter(as.Date(StartDate)=="2025-01-28") %>%
  filter(Progress != 14)

ideal_society_clean <- ideal_society_real_data %>%
  filter(Q127 != 5)

politicalorientation_data <- read.csv("C:\\Users\\Ben\\OneDrive\\Documents\\01Wake Forest\\2025SpringDataScience\\IdealSocietyData\\CopyMergefile.csv")

ideal_society_clean <- ideal_society_clean %>%
  left_join(politicalorientation_data, by = "PROLIFIC_PID")
  
ideal_society_clean <- ideal_society_clean %>%
  rename (poli_ori = PROLIFIC_Political.spectrum..us.)

ideal_society_clean <- ideal_society_clean %>%
  rename(LILH = Q7) %>%
  rename(LIHH = Q4) %>%
  rename(HILH = Q6) %>%
  rename(HIHH = Q5) %>%
  rename(LFLH = Q8) %>%
  rename(LFHH = Q124) %>%
  rename(HFLH = Q125) %>%
  rename(HFHH = Q126) %>%
  rename(LRLH = Q137) %>%
  rename(LRHH = Q139) %>%
  rename(HRLH = Q140) %>%
  rename(HRHH = Q141)

```

Next i want to remove variables I don't need for my analysis. That includes everything except for the society ratings ("LFLH" and so on) and the political ideology measures. I am taking out the immigration societies to simplify it further (we decided to not analyzing those societies).


```{r removingvariables}
ideal_society_simplified <- ideal_society_clean %>%
  select(LFLH,
         LFHH,
         HFLH,
         HFHH,
         LRLH,
         LRHH,
         HRLH,
         HRHH,
         Q2,
         Q26,
         Q27,
         Q28,
         poli_ori)
```

Now that it is just the simplified data, i want to zscore the political orientation measures. I am doing this to later average them and create a composite political orientation score. I need to z_score them because they are on different scales.

```{r zscore_polori}

ideal_society_simplified$Q2 <- as.numeric(ideal_society_simplified$Q2)
ideal_society_simplified$Q26 <- as.numeric(ideal_society_simplified$Q26)
ideal_society_simplified$Q27 <- as.numeric(ideal_society_simplified$Q27)
ideal_society_simplified$Q28 <- as.numeric(ideal_society_simplified$Q28)

ideal_society_simplified <- ideal_society_simplified %>%
  mutate(z_pol_ori = scale(Q2)) %>%
  mutate(z_pol_part_pref = scale(Q26)) %>%
  mutate(z_econ_descr = scale(Q27)) %>%
  mutate(z_soc_iss = scale(Q28))
  
```

Here i am making a new variable that is an average of the z scored political orientation variables. The goal of this is to use it as a predictor in the regression.

```{r aveargepoliori}
ideal_society_simplified <- ideal_society_simplified %>%  
  mutate(avrg_pol_score = rowMeans(select(ideal_society_simplified, z_pol_ori, z_pol_part_pref, z_econ_descr, z_soc_iss), na.rm = TRUE))

```

This averages the political orientation data and makes a new dataset.

```{r newdataset}

ideal_society_avrgpoli <- ideal_society_simplified %>%
  select(LFLH,
         LFHH,
         HFLH,
         HFHH,
         LRLH,
         LRHH,
         HRLH,
         HRHH,
         avrg_pol_score,
         poli_ori)
```


Here I am making a new variable that is the subject number. This is important for the regression analysis because it allows me to measure systematic differences between subjects. 

```{r newvariable_subject}
ideal_society_avrgpoli <- ideal_society_avrgpoli %>%
mutate(Subject = row_number())
```

Here i transformed the data from wide to long and re-coded the societies so that those with nontraditional family values (LF) are coded as 0 and traditional family values (HF) are coded as 1. I did the same for religion (LR = 0, HR = 1) and happiness (LH = 0, HH = 1). For the society types, 0 is coded as the one that might violate conservative values.Also conservatives are larger scores on the political orientation measure (i.e. 1 would be more conservative than -.4 which would be more liberal).

```{r pivotandrecode}
long_ideal_society <- ideal_society_avrgpoli %>%
 pivot_longer(
    cols = c(LFLH, LFHH, HFLH, HFHH, LRLH, LRHH, HRLH, HRHH),
    names_to = "Rating_Type",
    values_to = "Society_Rating"
  ) %>%
  mutate(
    Family_Values = ifelse(str_detect(Rating_Type, "LF"), 0, 
                           ifelse(str_detect(Rating_Type, "HF"), 1, NA)),
    Religion = ifelse(str_detect(Rating_Type, "LR"), 0, 
                      ifelse(str_detect(Rating_Type, "HR"), 1, NA)),
    Happiness = ifelse(str_detect(Rating_Type, "LH"), 0, 
                       ifelse(str_detect(Rating_Type, "HH"), 1, NA))) %>%
  select(Subject, Society_Rating, Family_Values, Religion, Happiness, avrg_pol_score, poli_ori)

```

I want to see how our collected political orientation data that is continuous compares to the self-reported Prolific data.
```{r poli_ori-check}

is_analysis <- long_ideal_society %>%
  mutate(
   dichotomized_poliori = ifelse(avrg_pol_score <= 0, "Liberal", "Conservative")
  )

```

Now I want to run code checking if this matches up with the Prolific data.
```{r poli_ori_comparison}

all(is_analysis$poli_ori == is_analysis$dichotomized_poliori, na.rm = TRUE)

###sum(..., na.rm = TRUE)
```


Here I am running a regression analysis on the data. I am predicting the society ratings based on the political orientation measure, happiness, and family values. I am also including a random effect for subject to account for systematic differences between subjects.

```{r regression}



long_ideal_society$Society_Rating <- as.numeric(long_ideal_society$Society_Rating)



family_mixedmodel <- lmer(Society_Rating ~ avrg_pol_score * Happiness * Family_Values + 
                          (1 | Subject), 
                          data = long_ideal_society %>% filter(!is.na(Family_Values)))
                            

summary(family_mixedmodel)

```
Here I am doing the same thing with the religion variable.

```{r religionmodel}


religion_mixedmodel <- lmer(Society_Rating ~ avrg_pol_score * Happiness * Religion + 
                          (1 | Subject), 
                          data = long_ideal_society %>% filter(!is.na(Religion)))
                            

summary(religion_mixedmodel)

```

```{r coefficients}
famval_model_coefs <- coef(summary(family_mixedmodel))
rel_model_coefs <- coef(summary(religion_mixedmodel))
```
