---
title: "Portfolio 4"
---

```{r load-packages, message=FALSE}
library(tidyverse) 
library(ggplot2)
library(viridis)

```



```{r getdata}
palmtrees <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-18/palmtrees.csv')

```
#exploratory graphs looking at the data

```{r summarystats}
palmtrees %>%
  ggplot(aes(x = max_stem_height_m)) +
  geom_histogram()
  
  palmtrees %>%
    ggplot(aes(x = max_stem_height_m)) +
  geom_histogram()+
  facet_wrap(~palm_subfamily)
  
  palmtrees %>%
    ggplot(aes(x = max_stem_height_m)) +
  geom_histogram()+
  facet_wrap(~fruit_shape)

   palmtrees %>%
    ggplot(aes(x = max_leaf_number)) +
  geom_histogram()+
  facet_wrap(~understorey_canopy)
   
palmtrees %>%
    ggplot(aes(x = max__blade__length_m)) +
  geom_histogram()+
  facet_wrap(~understorey_canopy)
  
```


Based on my exploratory results I want to create some good plots and statistical analyses to tell a story about how canopy and understory palms differ.

Quick googling showed me that canopy trees are those that are taller and receive more sunlight and wind compared to understory trees that receive less sunlight compared to the canopy trees.

#hypotheses

Some hypotheses that I have are:
1 Canopy trees are taller than understory trees (obvious but should be tested as a sanity check)
2 Fruit type will be different between groups (conspicuousness)
3  understorey trees might have more defenses compared to canopy trees (stem or leaves being armed)

#data_cleaning
```{r cleaning}
palmtrees_clean <- palmtrees %>%
  filter(understorey_canopy %in% c("canopy", "understorey"))
```

#h_1_taller
##t test

```{r taller}
palmtrees_clean %>%
  t.test(max_stem_height_m ~ understorey_canopy, data = .)
 
```

##violin_plot
```{r height_violin_plot}
palmtrees_clean %>%
  ggplot(aes(x = understorey_canopy, y = max_stem_height_m, fill = understorey_canopy)) +
  geom_violin() +
   scale_x_discrete(
    name   = "Understorey vs Canopy",          
    labels = c(
      canopy      = "Canopy",
      understorey = "Understorey")) +
  scale_fill_viridis_d(
    name   = "Type of Palm",
    labels = c(
      canopy      = "Canopy",
      understorey = "Understorey") ) +
  labs( y = "Max Stem Height (m)", 
        title = "Max Stem Height of Canopy vs Understorey Palms") +
  theme_classic()
  
```

##conclusion
Yes, the canopy palms appear to be much taller compared to the understorey palms. On average they are 14.78904 to 16.41920 (95% CI) meters taller.

#h_2_fruit
##testing_proportion
```{r conspiciousness}
tbl <- table(
  palmtrees_clean$understorey_canopy,
  palmtrees_clean$conspicuousness
)
prop.table(tbl, margin = 1)

prop.test(
  x = c(
    tbl["canopy",      "conspicuous"],    
    tbl["understorey", "conspicuous"]     
  ),
  n = c(
    sum(tbl["canopy", ]),                 
    sum(tbl["understorey", ])             
  )
)
```

##bar_plot
```{r fruit_bar_plot}
palmtrees_clean %>%
  filter(
    understorey_canopy %in% c("canopy", "understorey"),
    !is.na(conspicuousness)
  ) %>%
  ggplot(aes(x = understorey_canopy, fill = conspicuousness, na.rm = TRUE)) +
  geom_bar(position = "fill") +
  scale_x_discrete(
    name   = "Understorey vs Canopy",          
    labels = c(
      canopy      = "Canopy",
      understorey = "Understorey")) +
  scale_fill_viridis_d(
    name   = "Conspicuousness",
    labels = c(
      conspicuous      = "Conspicuous",
      inconspicuous = "Inconspicuous") ) +
  labs( y = "Proportion of Palms", 
        title = "Proportion of Fruit Conspicuousness in Canopy vs Understorey Palms") +
  theme_classic()
```


##conclusion

There is a statistically significant difference in the proportion of conspicuous fruit between canopy and understorey palms. Understorey palms are more likely to have conspiciuous fruit--meaning they are brighter in color--compared to canopy palms who are more likely to have cryptic fruit--fruit that have colors that are harder to see.

This might be because understorey palms may be in darker environments, meaning that they need to have brighter colored fruit in order to disperse their seeds where bright colors are not as adventatgeous in the canopy regions. 

#h_3_defense
##proportions
```{r palmdefensemeans}

#stem
tbl1 <- table(
  palmtrees_clean$understorey_canopy,
  palmtrees_clean$stem_armed)

prop.table(tbl1, margin = 1)

prop.test(
  x = c(
    tbl1["canopy",      "armed"],    
    tbl1["understorey", "armed"]     
  ),
  n = c(
    sum(tbl1["canopy", ]),           
    sum(tbl1["understorey", ])       
  )
)


#leaf
tbl2 <- table(
  palmtrees_clean$understorey_canopy,
  palmtrees_clean$leaves_armed)

prop.table(tbl2, margin = 1)

prop.test(
  x = c(
    tbl2["canopy",      "armed"],    
    tbl2["understorey", "armed"]     
  ),
  n = c(
    sum(tbl2["canopy", ]),           
    sum(tbl2["understorey", ])       
  )
)
  
```
##bar_plot
```{r armedbarplot}
palmtrees_clean %>%
  filter(
    understorey_canopy %in% c("canopy", "understorey"),
    !is.na(leaves_armed)
  ) %>%
  ggplot(aes(x = understorey_canopy, fill = leaves_armed, na.rm = TRUE)) +
  geom_bar(position = "fill") +
  scale_x_discrete(
    name   = "Understorey vs Canopy",          
    labels = c(
      canopy      = "Canopy",
      understorey = "Understorey")) +
  scale_fill_viridis_d(
    name   = "Leaves Armed",
    labels = c(
      armed      = "Armed",
      unarmed = "Unarmed") ) +
  labs( y = "Proportion of Palms", 
        title = "Proportion of Leaves Armed of Canopy vs Understorey Palms") +
  theme_classic()
```

##violinplot_armedxheight
```{r correlation}
palmtrees_clean %>%
  filter(
    leaves_armed %in% c("armed", "non-armed"),
    !is.na(max_stem_height_m)
  ) %>%
  
  mutate(
    leaves_armed = factor(leaves_armed, levels = c("non-armed","armed"))
  ) %>%
  ggplot(aes(
    x    = leaves_armed,
    y    = max_stem_height_m,
    fill = leaves_armed
  )) +
    geom_violin(trim = FALSE) +
    scale_x_discrete(
      name   = "Leaves Armed",
      labels = c("non-armed" = "Non-armed", "armed" = "Armed")
    ) +
    scale_fill_viridis_d(
      name   = "Leaf Armour",
      labels = c("Non-armed", "Armed")
    ) +
    labs(
      y     = "Max Stem Height (m)",
      title = "Distribution of Stem Height by Leaf Armament"
    ) +
    theme_classic()
```

##conclusion

understorey palms seem to be more likely to have stem armour (p = .005) and leaf armour (p < .001). Leaf armour also is more common for shorter palms, supporting this hypothesis. 