---
title: "Portfolio 4"
---

```{r load-packages, message=FALSE}
library(tidyverse) 
library(ggplot2)
library(viridis)

```



```{r getdata}
palmtrees <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-18/palmtrees.csv')

```
exploratory graphs looking at the data

```{r summarystats}
palmtrees %>%
  ggplot(aes(x = max_stem_height_m)) +
  geom_histogram()
  
  palmtrees %>%
    ggplot(aes(x = max_stem_height_m)) +
  geom_histogram()+
  facet_wrap(~palm_subfamily)
  
  palmtrees %>%
    ggplot(aes(x = max_stem_height_m)) +
  geom_histogram()+
  facet_wrap(~fruit_shape)

   palmtrees %>%
    ggplot(aes(x = max_leaf_number)) +
  geom_histogram()+
  facet_wrap(~understorey_canopy)
   
palmtrees %>%
    ggplot(aes(x = max__blade__length_m)) +
  geom_histogram()+
  facet_wrap(~understorey_canopy)
  
```


Based on my exploratory results I want to create some good plot to tell a story about how canopy and understory palms differ.

Quick googling showed me that canopy trees are those that are taller and receive more sunlight and wind compared to understory trees that receive less sunlight compared to the canopy trees.

#hypothesis

Some hypotheses that I have are:
1 Canopy trees are taller than understory trees 
2 Understory trees will have larger leaves in order to get more sun (not blade length)
3 *exploratory* understorey trees might have more defenses compared to canopy trees (stem or leaves being armed)

#data_cleaning
```{r cleaning}
palmtrees_clean <- palmtrees %>%
  filter(understorey_canopy %in% c("canopy", "understorey"))
```

#h_1_taller
##t test

```{r taller}
palmtrees_clean %>%
  t.test(max_stem_height_m ~ understorey_canopy, data = .)
 
```

##violin_plot
```{r height_violin_plot}
palmtrees_clean %>%
  ggplot(aes(x = understorey_canopy, y = max_stem_height_m, fill = understorey_canopy)) +
  geom_violin() +
   scale_x_discrete(
    name   = "Understorey vs Canopy",          
    labels = c(
      canopy      = "Canopy",
      understorey = "Understorey")) +
  scale_fill_viridis_d(
    name   = "Type of Palm",
    labels = c(
      canopy      = "Canopy",
      understorey = "Understorey") ) +
  labs( y = "Max Stem Height (m)", 
        title = "Max Stem Height of Canopy vs Understorey Palms") +
  theme_classic()
  
```
