---
title: "Portfolio 5"
---

```{r}
library(tidytuesdayR)
library(tidyverse)
library(ggplot2)

```

```{r}
tuesdata <- tidytuesdayR::tt_load('2024-11-19')
episode_metrics <- tuesdata$episode_metrics

```
What season has the most questions?
```{r question}
ggplot(episode_metrics, aes(season, question_ratio)) +
  geom_col() +
  labs(title = "Season with the most questions",
       x = "Season",
       y = "Number of questions")
```

correlation of questions and exlamations
```{r corr}
ggplot(episode_metrics, aes(question_ratio, exclamation_ratio)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Correlation of questions and exclamations",
       x = "Number of questions",
       y = "Number of exclamations")+
  theme_classic()
```
dialogue density
```{r dialogue density}
ggplot(episode_metrics, aes(x = season, y = dialogue_density)) +
  geom_point() +
  labs(title = "Dialogue density",
       x = "Season",
       y = "Dialogue density")
```

I predict that an absolute value of sentiment variance (I realized afterwards that although AFINN scores can be negative, none of the episodes had a negative score so this turning this into an absolute value was unnecessary) will be positively correlated with ratio of exclamations
```{r AFINN-exclam}
library(viridis)


episode_metrics <- episode_metrics %>%
  mutate(AFINN_abs = abs(sentiment_variance))

ggplot(episode_metrics, aes(AFINN_abs, exclamation_ratio)) +
  geom_point(aes(color = season)) +
  geom_smooth(method = "lm", color = "black") +
  labs(title = "Correlation of AFINN Scores of and Ratio of Exclamations in Bobs Burger Episodes",
       x = "AFINN Score",
       y = "Ratio of Exclamations",
       color = "Season"
      )+
  theme_classic()+ 
  scale_colour_viridis_c(option = "plasma")
```

because "Linn" is annoying, i have to check if this is significant
```{r Linn}
cor.test(episode_metrics$AFINN_abs, episode_metrics$exclamation_ratio, method = "pearson")

```
Here i want to check to make sure this isn't driven primarily by only the negative or positive scores
