---
title: "Portfolio 5"
---

```{r}
library(tidytuesdayR)
library(tidyverse)
library(ggplot2)

```

```{r}
tuesdata <- tidytuesdayR::tt_load('2024-11-19')
episode_metrics <- tuesdata$episode_metrics

```
What season has the most questions?
```{r question}
episode_metrics %>%
  group_by(season) %>%
  summarise(average_qr = mean(question_ratio, na.rm = TRUE)) %>%
  ggplot(aes
         (x = season, y = average_qr)) +
  geom_col(fill = "#B28142") +
  labs(title = "Average Question Ratio per Season",
       x = "Season",
       y = "Average question ratio") +
  theme_classic() 
```

I choose the color to matach the color of burger buns!

I am predicting that the ratio of questions and exlamations in the dialogue will correlate significantly.
```{r corr}
ggplot(episode_metrics, aes(question_ratio, exclamation_ratio)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Correlation of questions and exclamations",
       x = "Ratio of Questions",
       y = "Ratio of Exclamations")+
  theme_classic()
```

```{r corr-test}
cor.test(episode_metrics$question_ratio, episode_metrics$exclamation_ratio, method = "pearson")

```


dialogue density ***exploratory**
```{r dialogue density}
ggplot(episode_metrics, aes(x = season, y = dialogue_density)) +
  geom_point() +
  labs(title = "Dialogue density",
       x = "Season",
       y = "Dialogue density")
```

I predict that an absolute value of sentiment variance (I realized afterwards that although AFINN scores can be negative, none of the episodes had a negative score so this turning this into an absolute value was unnecessary) will be positively correlated with ratio of exclamations
```{r AFINN-exclam}
library(viridis)


episode_metrics <- episode_metrics %>%
  mutate(AFINN_abs = abs(sentiment_variance))

ggplot(episode_metrics, aes(AFINN_abs, exclamation_ratio)) +
  geom_point(aes(color = season)) +
  geom_smooth(method = "lm", color = "black") +
  labs(title = "Correlation of AFINN Scores of and Ratio of Exclamations in Bobs Burger Episodes",
       x = "AFINN Score",
       y = "Ratio of Exclamations",
       color = "Season"
      )+
  theme_classic()+ 
  scale_colour_viridis_c(option = "plasma")
```


```{r Linnmademedothis}
cor.test(episode_metrics$AFINN_abs, episode_metrics$exclamation_ratio, method = "pearson")

```
