---
title: "Portfolio 6"
---

This project is to remake my graphs from line graphs to bar graphs 

Here is what i had before 
#graphs that need to be fixed

```{r libraries}
library(tidyverse)
library(ggplot2)
library(extrafont)
library(gridExtra)

```

```{r df_3}
means <- c(9.17, 8.59, 5.67, 6.56, 3.11, 3.19, 5.65, 4.23, 8.70, 9.86, 3.33, 4.15, 7.90, 5.86, 3.16, 2.15)
se <- c(.112, .104, .176, .131, .149, .119, .169, .131, .169, .086, .177, .121, .150, .125, .173, .091)
happiness <- c("high happiness", "high happiness", "high happiness", "high happiness", "low happiness", "low happiness", "low happiness", "low happiness", "high happiness", "high happiness", "low happiness", "low happiness", "high happiness", "high happiness", "low happiness", "low happiness")
political_orientation <- c("cons", "lib", "cons", "lib", "cons", "lib", "cons", "lib", "cons", "lib", "cons", "lib", "cons", "lib", "cons", "lib")
society_type <- c("hmhh", "hmhh", "lmhh", "lmhh", "lmlh", "lmlh","hmlh", "hmlh", "hehh", "hehh", "helh", "helh", "lehh", "lehh", "lelh", "lelh")
df_3 <- data.frame(means, se, happiness, political_orientation, society_type)


df_3 <- df_3 %>%
  mutate(
    happiness = fct_relevel(happiness, "low happiness", "high happiness"))
```


```{r fig_3_prep}
df_3 <- df_3 %>%
  mutate(society_type = case_when(
    str_detect(society_type, "lm") ~ "Low Meritocracy/High Equality",
    str_detect(society_type, "hm") ~ "High Meritocracy/Low Equality",
    str_detect(society_type, "le") ~ "Low Equality",
    str_detect(society_type, "he") ~ "High Equality",
    TRUE ~ society_type  # Keeps the original value if none of the conditions match
  ))
df_3<- df_3 %>%
  mutate(happiness_numeric = ifelse(happiness == "low happiness", 0, 1))

df_3_meri <- df_3 %>%
  filter(society_type %in% c("Low Meritocracy/High Equality", "High Meritocracy/Low Equality"))

# Filter for "Low Equality" or "High Equality"
df_3_eq <- df_3 %>%
  filter(society_type %in% c("Low Equality", "High Equality"))
```

```{r fig}
df_3_meri_plot <- ggplot(df_3_meri, 
  aes(x = happiness_numeric + ifelse(society_type == "High Meritocracy/Low Equality", 0.1, -0.1), 
      y = means, color = political_orientation, 
      group = interaction(political_orientation, society_type), 
      linetype = society_type)) +  
  geom_point(shape = 20, size = 5) +  # Points for means
  geom_line(linewidth = .5) + 
  geom_errorbar(aes(ymin = means - se, ymax = means + se), width = 0.05, linewidth = 0.5) +# Connecting means
  scale_color_manual(
    values = c("cons" = "red", "lib" = "blue"), 
    name = "Political Orientation", 
    labels = c("cons" = "Conservatives", "lib" = "Liberals")
  ) + 
  scale_linetype_manual(values = c("Low Meritocracy/High Equality" = "solid", "High Meritocracy/Low Equality" = "dashed")) +  # Line types for violations
  scale_x_continuous(
    breaks = c(0, 1),
    labels = c("Low Happiness", "High Happiness")
  ) +
  scale_y_continuous(breaks = seq(1, 11, by = 1), limits = c(1, 11)) + 
  labs(
    title = NULL,
    x = NULL,
    y = "Ideal Society Rating",
    color = "Political Orientation",
    linetype = "Society Type"
  ) + 
  theme_classic() + 
  theme(
    text = element_text(family = "Times New Roman", size = 12),
    legend.position = "right",
    axis.text.x = element_text(size = 12),
    strip.background = element_blank(),
    strip.text = element_text(size = 12)
  )

print(df_3_meri_plot)
```

Here is what did to fix them 

```{r df_stdy3}
means <- c(9.17, 8.59, 5.67, 6.56, 3.11, 3.19, 5.65, 4.23, 8.70, 9.86, 3.33, 4.15, 7.90, 5.86, 3.16, 2.15)
se <- c(.112, .104, .176, .131, .149, .119, .169, .131, .169, .086, .177, .121, .150, .125, .173, .091)
happiness <- c("high happiness", "high happiness", "high happiness", "high happiness", "low happiness", "low happiness", "low happiness", "low happiness", "high happiness", "high happiness", "low happiness", "low happiness", "high happiness", "high happiness", "low happiness", "low happiness")
political_orientation <- c("cons", "lib", "cons", "lib", "cons", "lib", "cons", "lib", "cons", "lib", "cons", "lib", "cons", "lib", "cons", "lib")
society_type <- c("hmhh", "hmhh", "lmhh", "lmhh", "lmlh", "lmlh","hmlh", "hmlh", "hehh", "hehh", "helh", "helh", "lehh", "lehh", "lelh", "lelh")
df_stdy3 <- data.frame(means, se, happiness, political_orientation, society_type)


df_stdy3 <- df_stdy3 %>%
  mutate(
    happiness = fct_relevel(happiness, "low happiness", "high happiness"))
```

```{r fig1prep}
df_stdy3 <- df_stdy3 %>%
  mutate(society_type = case_when(
    str_detect(society_type, "lm") ~ "Low Meritocracy/High Equality",
    str_detect(society_type, "hm") ~ "High Meritocracy/Low Equality",
    str_detect(society_type, "le") ~ "Low Equality",
    str_detect(society_type, "he") ~ "High Equality",
    TRUE ~ society_type  # Keeps the original value if none of the conditions match
  ))
df_stdy3<- df_stdy3 %>%
  mutate(happiness_numeric = ifelse(happiness == "low happiness", 0, 1))

df_stdy3_meri <- df_stdy3 %>%
  filter(society_type %in% c("Low Meritocracy/High Equality", "High Meritocracy/Low Equality"))

# Filter for "Low Equality" or "High Equality"
df_stdy3_eq <- df_stdy3 %>%
  filter(society_type %in% c("Low Equality", "High Equality"))

df_stdy3_meri$society_type <- factor(
  df_stdy3_meri$society_type,
  levels = c("Low Meritocracy/High Equality", "High Meritocracy/Low Equality"))

df_stdy3_meri$happiness_numeric <- factor(df_stdy3_meri$happiness_numeric, 
                                           levels = c(0, 1), 
                                           labels = c("Low", "High"))
df_stdy3_eq$happiness_numeric <- factor(df_stdy3_eq$happiness_numeric, 
                                           levels = c(0, 1), 
                                           labels = c("Low", "High"))

```

##stdy3
###fig1_stdy3A




```{r fig1_stdy3A}
df_stdy3_meri_plot_bar <- ggplot(df_stdy3_meri, aes(x = happiness_numeric, y = means, fill = society_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = means - se, ymax = means + se), 
                width = 0.2, position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("High Meritocracy/Low Equality" = "gray50", 
                               "Low Meritocracy/High Equality" = "gray80")) +  # Bar fill colors
  scale_y_continuous(
    breaks = seq(1, 11, by = 1)
  ) + 
  labs(
    title = NULL,
    x = "Happiness",
    y = "Mean Ideal Society Rating",
    fill = "Society Type:"
  ) + 
  theme_classic() + 
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 15)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    strip.background = element_blank(),
    strip.text = element_text(size = 10),
    legend.position = "bottom",  # Legend at the bottom
    legend.key.size = unit(0.5, "cm"),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10)
  ) +
   facet_wrap(~political_orientation, scales = "fixed", labeller = labeller(political_orientation = c("cons" = "Conservatives", "lib" = "Liberals"))) +
  coord_cartesian(ylim = c(1, 11))

print(df_stdy3_meri_plot_bar)



```
###fig2_stdy3B

```{r fig2_stdy3B}
df_stdy3_eq_plot <- ggplot(df_stdy3_eq, aes(x = happiness_numeric, y = means, fill = society_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = means - se, ymax = means + se), 
                width = 0.2, position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("Low Equality" = "gray50", 
                               "High Equality" = "gray80")) +  # Bar fill colors
  scale_y_continuous(
    breaks = seq(1, 11, by = 1)
  ) + 
  labs(
    title = NULL,
    x = "Happiness",
    y = "Mean Ideal Society Rating",
    fill = "Society Type:"
  ) + 
  theme_classic() + 
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 15)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    strip.background = element_blank(),
    strip.text = element_text(size = 10),
    legend.position = "bottom",  # Legend at the bottom
    legend.key.size = unit(0.5, "cm"),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10)
  ) +
   facet_wrap(~political_orientation, scales = "fixed", labeller = labeller(political_orientation = c("cons" = "Conservatives", "lib" = "Liberals"))) +
  coord_cartesian(ylim = c(1, 11))

print(df_stdy3_eq_plot)



```

#hypothesis



I will make two graphs. The first will depict the more intuitive results (two main effects--John's question from the defense). The next graph will show our hypothesis.



```{r df_intuit}
isr <- c(5,3,10,8)
happiness <- c("Low Happiness", "Low Happiness", "High Happiness", "High Happiness")
violation <- c("No Violation", "Violation", "No Violation", "Violation")
intuit <- data.frame(isr, happiness, violation)

intuit <- intuit %>%
  mutate(
    happiness = fct_relevel(happiness, "Low Happiness", "High Happiness")
  )
```

###fig4_hyp_intuit
```{r fig_4}
fig4_intuit <- ggplot(intuit, aes(x = happiness, y = isr, fill = violation)) +
  geom_bar(stat = "identity", position = "dodge")+
  scale_fill_manual(values = c("No Violation" = "gray80", "Violation" = "gray50")) +  # Fill colors for bars
  scale_y_continuous(breaks = 1:11) +  # Y-axis scale
  coord_cartesian(ylim = c(1, 11)) +
  theme_classic() +
  theme( 
    text = element_text(family = "Times New Roman"),
    axis.text.x = element_text(family = "Times New Roman", size = 12),    # Match the font of the x-axis labels
    axis.title.y = element_text(family = "Times New Roman", size = 12)
  ) +
  labs(
    title = NULL,
    x = NULL,
    y = "Ideal Society Rating",
    fill = "Violation"
  )

print(fig4_intuit)

```


```{r fig3_predicted}
isr <- c(5,3,10,4)
happiness <- c("Low Happiness", "Low Happiness", "High Happiness", "High Happiness")
violation <- c("No Violation", "Violation", "No Violation", "Violation")
predicted <- data.frame(isr, happiness, violation)

predicted <- predicted %>%
  mutate(
    happiness = fct_relevel(happiness, "Low Happiness", "High Happiness")
  )
```

###fig3_hyp_prediction
```{r fig_3}
fig3_predicted <-ggplot(predicted, aes(x = happiness, y = isr, fill = violation)) +
  geom_bar(stat = "identity", position = "dodge")+
  scale_fill_manual(values = c("No Violation" = "gray80", "Violation" = "gray50")) +
   scale_y_continuous(breaks = 1:11) + 
    coord_cartesian(ylim = c(1, 11)) +
  theme_classic()+  
  theme( 
    text = element_text(family = "Times New Roman"),
    axis.text.x = element_text(family = "Times New Roman", size = 12),    # Match the font of the x-axis labels
    axis.title.y = element_text(family = "Times New Roman", size = 12)
  ) +
  labs(
    title = NULL,
    x = NULL,
    y = "Ideal Society Rating",
    fill = "Violation"
  )


print(fig3_predicted)

```

##three-way


```{r dataforthreeway}
ideal_society_ratings <- c(4, 2, 8, 6, 2, 3, 3, 7)
happiness <- c("low happiness", "low happiness", "high happiness", "high happiness", "low happiness", "low happiness", "high happiness", "high happiness")
political_orientation <- c("cons", "lib", "cons", "lib", "cons", "lib", "cons", "lib")
society_type <- c("no violation", "no violation", "no violation", "no violation", "violation", "violation", "violation", "violation")
df_hyp_tw <- data.frame(ideal_society_ratings, happiness, political_orientation, society_type)


df_hyp_tw <- df_hyp_tw %>%
  mutate(
    happiness = fct_relevel(happiness, "low happiness", "high happiness"))
df_hyp_tw$happiness <- factor(df_hyp_tw$happiness, levels = c("low happiness", "high happiness"), labels = c("Low Happiness", "High Happiness"))
df_hyp_tw$society_type <- factor(df_hyp_tw$society_type, 
                                 levels = c("no violation", "violation"), 
                                 labels = c("Conservative Values Are Not Violated", "Conservative Values Are Violated"))
```

###fig5_hyp_three-way
```{r fig5_hyp_three-way}
fig5_tw_predicted <- ggplot(df_hyp_tw, aes(x = happiness, y = ideal_society_ratings, fill = society_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("Conservative Values Are Not Violated" = "gray80", "Conservative Values Are Violated" = "gray50")) +  # Bar fill colors
  scale_y_continuous(breaks = seq(1, 11, by = 1)) + 
  labs(
    title = NULL,
    x = "Happiness",
    y = "Ideal Society Rating",
    fill = "Society Type:"
  ) + 
  theme_classic() + 
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 15)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    strip.background = element_blank(),
    strip.text = element_text(size = 10),
    legend.position = "bottom",  # Legend at the bottom
    legend.key.size = unit(0.5, "cm"),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10)
  ) +
  facet_wrap(~political_orientation, scales = "fixed", labeller = labeller(political_orientation = c("cons" = "Conservatives", "lib" = "Liberals"))) +
  coord_cartesian(ylim = c(1, 11))

print(fig5_tw_predicted)

```

#RESULTS
#data_means_se


```{r data}

df_4 <- tibble(
  poli_ori = rep(c("Conservative", "Liberal"), each = 8),
  society_type = rep(c("LF", "LF", "HF", "HF", "LR", "LR", "HR", "HR"), 2),
  happiness = rep(c("low happiness", "high happiness"), times = 8),
  mean = c(
    2.627706, 6.341991, 4.904762, 9.835498, 2.816479, 6.996255, 4.501873, 9.591760,
    3.570370, 9.462963, 2.714815, 7.644444, 3.208696, 9.786957, 2.426087, 7.286957
  ),
  se = c(
    0.14978388, 0.17792659, 0.18200138, 0.09402925, 0.13939406, 0.18727943, 0.16487943, 0.10800660,
    0.13674704, 0.10169496, 0.12955408, 0.13962178, 0.14105724, 0.09826295, 0.11263820, 0.18399305
  )
) |>
  mutate(
    happiness = fct_relevel(happiness, "low happiness", "high happiness"),
    happiness_numeric = if_else(happiness == "low happiness", 0, 1)
  )

# Filter subsets
famval <- df_4 %>% filter(str_detect(society_type, "F"))
Christ <- df_4 %>% filter(str_detect(society_type, "R"))
```

##fig6_fam_means

```{r fig6_famval_means}
famval_means <- ggplot(famval, aes(x = happiness, y = mean, fill = society_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(
    aes(ymin = mean - se, ymax = mean + se),
    width = 0.2, position = position_dodge(width = 0.9)) +
  scale_fill_manual(
    values = c("LF" = "gray50", "HF" = "gray80"),
    labels = c(
      "LF" = "Nontraditional Family Values",
      "HF" = "Traditional Family Values"
    )
  ) +
  
  scale_y_continuous(
    breaks = seq(1, 11, by = 1)
  ) +
  
  scale_x_discrete(
    labels = c("low happiness" = "Low", "high happiness" = "High")
  ) +
  
  labs(
    title = NULL,
    x = "Happiness",
    y = "Mean Ideal Society Rating",
    fill = "Society Type:"
  ) +
  
  theme_classic() +
  
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 15)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    strip.background = element_blank(),
    strip.text = element_text(size = 10),
    legend.position = "bottom",
    legend.key.size = unit(0.5, "cm"),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10)
  ) +
  
  facet_wrap(
    ~poli_ori,
    labeller = labeller(political_orientation = c("cons" = "Conservatives", "lib" = "Liberals"))
  ) +
  
  coord_cartesian(ylim = c(1, 11))

# Print and save
print(famval_means)

```

##fig12_chris_means
```{r fig12_christ_means}
christ_means <- ggplot(Christ, aes(x = happiness, y = mean, fill = society_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(
    aes(ymin = mean - se, ymax = mean + se),
    width = 0.2, position = position_dodge(width = 0.9)) +
  scale_fill_manual(
    values = c("LR" = "gray50", "HR" = "gray80"),
    labels = c(
      "LR" = "Christianity is Unimportant",
      "HR" = "Christianity is Important"
    )
  ) +
  
  scale_y_continuous(
    breaks = seq(1, 11, by = 1)
  ) +
  
  scale_x_discrete(
    labels = c("low happiness" = "Low", "high happiness" = "High")
  ) +
  
  labs(
    title = NULL,
    x = "Happiness",
    y = "Mean Ideal Society Rating",
    fill = "Society Type:"
  ) +
  
  theme_classic() +
  
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 15)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    strip.background = element_blank(),
    strip.text = element_text(size = 10),
    legend.position = "bottom",
    legend.key.size = unit(0.5, "cm"),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10)
  ) +
  
  facet_wrap(
    ~poli_ori,
    labeller = labeller(political_orientation = c("cons" = "Conservatives", "lib" = "Liberals"))
  ) +
  
  coord_cartesian(ylim = c(1, 11))

# Print and save
print(christ_means)

```

#MainEffects
###fam_data
```{r main_effects}
fam_main_effects <- tribble(
  ~isr_mean, ~main_effect, ~se,
  5.84, "pol0", 0.070,
  5.93, "pol1", 0.076,
  5.50, "lf", 0.075,
  6.27, "hf", 0.073,
  3.45, "lh", 0.090,
  8.32, "hh", 0.067
)

fam_main_effects$facet_group <- rep(c("pol", "society", "happiness"), each = 2)
df_main_effects <- fam_main_effects %>%
  mutate(main_effect = factor(main_effect, levels = c("lh", "hh", "pol0", "pol1", "lf", "hf")))

fam_maineffects_long <- df_main_effects %>%
  mutate(factor_type = case_when(
    main_effect %in% c("pol0", "pol1") ~ "Political Orientation",
    main_effect %in% c("lf", "hf") ~ "Family Values",
    main_effect %in% c("lh", "hh") ~ "Happiness Level"
  ))


```

##fig7_fam_ME
```{r famval_main_effects}
fam_maineffect <- ggplot(fam_maineffects_long, aes(x = main_effect, y = isr_mean, fill = main_effect)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = isr_mean - se, ymax = isr_mean + se), 
                width = 0.2, position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("pol0" = "gray50", "pol1" = "gray80", 
                               "lf" = "gray50", "hf" = "gray80", 
                               "lh" = "gray50", "hh" = "gray80")) +
  scale_x_discrete(labels = c("pol0" = "Liberal", "pol1" = "Conservative", 
                              "lf" = "Nontraditional", "hf" = "Traditional", 
                              "lh" = "Low", "hh" = "High")) +
  scale_y_continuous(breaks = 1:11) +
  coord_cartesian(ylim = c(1, 11)) +
  labs(
    title = NULL,
    x = NULL,
    y = "Mean Ideal Society Rating",
    fill = NULL
  ) +
  theme_classic() +
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 15)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    legend.position = "none",
    strip.text = element_text(size = 12)  # Add size for facet labels
  ) +
  facet_wrap(~factor_type, scales = "free")  # Facet by factor_type

# Print the combined plot
print(fam_maineffect)

```

##christ_data
```{r christ_ME_data}
christ_main_effects <- tribble(
  ~isr_mean, ~main_effect, ~se,
  5.68, "pol0", 0.075,
  5.98, "pol1", 0.070,
  5.70, "lr", 0.079,
  5.95, "hr", 0.078,
  3.24, "lh", 0.084,
  8.42, "hh", 0.071
)

christ_main_effects$facet_group <- rep(c("pol", "society", "happiness"), each = 2)

# Fix typo in the `mutate()` function and correct variable name
christ_main_effects <- christ_main_effects %>%
  mutate(main_effect = factor(main_effect, levels = c("lh", "hh", "pol0", "pol1", "lr", "hr")))

# Convert data into long format and add a factor type
christ_maineffects_long <- christ_main_effects %>%
  mutate(factor_type = case_when(
    main_effect %in% c("pol0", "pol1") ~ "Political Orientation",
    main_effect %in% c("lr", "hr") ~ "Christianity",
    main_effect %in% c("lh", "hh") ~ "Happiness Level"
  ))

```

##fig13_christ_ME

```{r fig13_christ_maineffect}
christ_maineffect <- ggplot(christ_maineffects_long, aes(x = main_effect, y = isr_mean, fill = main_effect)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = isr_mean - se, ymax = isr_mean + se), 
                width = 0.2, position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("pol0" = "gray50", "pol1" = "gray80", 
                               "lr" = "gray50", "hr" = "gray80", 
                               "lh" = "gray50", "hh" = "gray80")) +
  scale_x_discrete(labels = c("pol0" = "Liberal", "pol1" = "Conservative", 
                              "lr" = "Unimportant", "hr" = "Important", 
                              "lh" = "Low", "hh" = "High")) +
  scale_y_continuous(breaks = 1:11) +
  coord_cartesian(ylim = c(1, 11)) +
  labs(
    title = NULL,
    x = NULL,
    y = "Mean Ideal Society Rating",
    fill = NULL
  ) +
  theme_classic() +
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 15)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    legend.position = "none",
    strip.text = element_text(size = 12)  # Add size for facet labels
  ) +
  facet_wrap(~factor_type, scales = "free")  # Facet by factor_type

# Print the combined plot
print(christ_maineffect)


```
#two_way
##famval

```{r famval_tw_data}
fam_tw_df <- tribble(
 ~poli_ori, ~Happiness, ~emmean, ~SE,
  "Liberal", "Low", 3.14, 0.1220, 
  "Conservative", "Low", 3.77, 0.1320,
  "Liberal", "High", 8.55, 0.0907,
  "Conservative", "High", 8.09, 0.0981
)

fam_tw_df$Happiness <- factor(fam_tw_df$Happiness, levels = c("Low", "High"))

```

##fig8_famPxH
```{r fig8_tw_pxh}
fam_tw_plot <- ggplot(fam_tw_df, aes(x = Happiness, y = emmean, fill = Happiness)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), 
                width = 0.2, position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("Low" = "gray50", "High" = "gray80")) +  # Custom colors for Low and High Happiness
  scale_y_continuous(breaks = seq(1, 11, by = 1)) +
  coord_cartesian(ylim = c(1, 11)) +
  labs(
    title = NULL,
    x = "Happiness Level",
    y = "Mean Ideal Society Rating",
    fill = "Happiness"
  ) +
  theme_classic() +
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 15)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    legend.position = "none"  # Remove legend if not needed
  ) +
  facet_wrap(~poli_ori)  # Facet by poli_ori (Liberal vs Conservative)

# Print the plot
print(fam_tw_plot)

```

```{r fig9_data}
fam_tw_ps <- tribble(
 ~poli_ori, ~society_type, ~emmean, ~SE,
  "Liberal", "Nontraditional", 6.52, 0.10, 
  "Conservative", "Nontraditional", 4.48, 0.11,
  "Liberal", "Traditional", 5.18, 0.099,
  "Conservative", "Traditional", 7.37, 0.11
)
fam_tw_ps$society_type <- factor(fam_tw_ps$society_type, levels = c("Traditional", "Nontraditional"))
```

##fig9_famPxS
```{r fig9_fam_polixsociet}
fam_tw_ps <- ggplot(fam_tw_ps, aes(x = society_type, y = emmean, fill = society_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), 
                width = 0.2, position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("Nontraditional" = "gray50", "Traditional" = "gray80")) +  # Custom colors for Low and High Happiness
  scale_y_continuous(breaks = seq(1, 11, by = 1)) +
  coord_cartesian(ylim = c(1, 11)) +
  labs(
    title = NULL,
    x = "Family Values",
    y = "Mean Ideal Society Rating",
    fill = "Society Type"
  ) +
  theme_classic() +
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 15)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    legend.position = "none"  # Remove legend if not needed
  ) +
  facet_wrap(~poli_ori)  # Facet by poli_ori (Liberal vs Conservative)

# Print the plot
print(fam_tw_ps)
# Save the plot


```

```{r fig10data}
fam_tw_hs <- tribble(
 ~Happiness, ~society_type, ~emmean, ~SE,
  "Low", "Nontraditional", 3.10, 0.10, 
  "High", "Nontraditional", 7.9, 0.10,
  "Low", "Traditional", 3.81, 0.11,
  "High", "Traditional", 8.74, 0.087
)
fam_tw_hs$Happiness <- factor(fam_tw_hs$Happiness, levels = c("Low", "High"))
```

##fig10_famHxS
```{r fig10_fam_hapxsociety}
fam_tw_hs_plot <- ggplot(fam_tw_hs, aes(x = Happiness, y = emmean, fill = Happiness)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), 
                width = 0.2, position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("Low" = "gray50", "High" = "gray80")) +  # Custom fill
  scale_y_continuous(breaks = seq(1, 11, by = 1)) +
  coord_cartesian(ylim = c(1, 11)) +
  labs(
    title = NULL,
    x = "Happiness",
    y = "Mean Ideal Society Rating",
    fill = "Happiness"
  ) +
  theme_classic() +
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 15)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    legend.position = "none"
  ) +
  facet_wrap(~society_type)  # ✅ Facet by society_type

# Print the plot
print(fam_tw_hs_plot)

```


##christianity
```{r christ_tw_poli-happiness_data}
christ_tw_ph <- tribble(
 ~political_orientation, ~Happiness, ~emmean, ~SE,
  "Liberal", "Low", 2.82, 0.12, 
  "Conservative", "Low", 3.66, 0.12,
  "Liberal", "High", 8.54, 0.10,
  "Conservative", "High", 8.29, 0.10
)

christ_tw_ph$Happiness <- factor(christ_tw_ph$Happiness, levels = c("Low", "High"))
```


##fig14_ch_PxH
```{r fig14_christian_tw_poli_happiness}
christ_tw_ph_plot <- ggplot(christ_tw_ph, aes(x = Happiness, y = emmean, fill = Happiness)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), 
                width = 0.2, position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("Low" = "gray50", "High" = "gray80")) +  # Custom fill
  scale_y_continuous(breaks = seq(1, 11, by = 1)) +
  coord_cartesian(ylim = c(1, 11)) +
  labs(
    title = NULL,
    x = "Happiness",
    y = "Mean Ideal Society Rating",
    fill = "Happiness"
  ) +
  theme_classic() +
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 15)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    legend.position = "none"
  ) +
  facet_wrap(~political_orientation)
print(christ_tw_ph_plot)

```

```{r christ_tw_PxS_data}
christ_tw_ps <- tribble(
 ~political_orientation, ~society_type, ~emmean, ~SE,
  "Liberal", "Unimportant", 6.5, 0.12, 
  "Conservative", "Unimportant", 4.91, 0.11,
  "Liberal", "Important", 4.86, 0.11,
  "Conservative", "Important", 7.05, 0.11)
```

##fig15_ch_PxS
```{r fig15_christian_politicalorientation_societytype}
christ_tw_ps_plot <- ggplot(christ_tw_ps, aes(x = society_type, y = emmean, fill = society_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), 
                width = 0.2, position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("Unimportant" = "gray50", "Important" = "gray80")) +  # Custom fill
  scale_y_continuous(breaks = seq(1, 11, by = 1)) +
  coord_cartesian(ylim = c(1, 11)) +
  labs(
    title = NULL,
    x = "Christianity",
    y = "Mean Ideal Society Rating",
    fill = "Christianity"
  ) +
  theme_classic() +
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 15)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    legend.position = "none"
  )    +
      facet_wrap(~political_orientation)
    print(christ_tw_ps_plot)
    
```
  
```{r christian_tw_society_happiness_data}
christ_tw_hs <- tribble(
 ~Happiness, ~society_type, ~emmean, ~SE,
  "Low", "Unimportant", 3.01, 0.10, 
  "High", "Unimportant", 8.39, 0.11,
  "Low", "Important", 3.46, 0.10,
  "High", "Important", 8.44, 0.10)
christ_tw_hs$Happiness <- factor(christ_tw_hs$Happiness, levels = c("Low", "High"))
```


##fig16_ch_HxS
```{r fig16_christian_tw_happiness_societytype}
christ_tw_hs_plot <- ggplot(christ_tw_hs, aes(x = Happiness, y = emmean, fill = Happiness)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), 
                width = 0.2, position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("Low" = "gray50", "High" = "gray80")) +  # Custom fill
  scale_y_continuous(breaks = seq(1, 11, by = 1)) +
  coord_cartesian(ylim = c(1, 11)) +
  labs(
    title = NULL,
    x = "Happiness",
    y = "Mean Ideal Society Rating",
    fill = "Happiness"
  ) +
  theme_classic() +
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 15)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    legend.position = "none"
  ) +
  facet_wrap(~society_type)
print(christ_tw_hs_plot)

```

#three-way
```{r famval_3w_data}
famval_3w_df <- tribble(
  ~poli_ori,      ~society_type, ~Happiness,         ~mean,       ~se,
  "Conservative", "Nontraditional",          "Low",     2.627706,    0.14978388,
  "Conservative", "Nontraditional",          "High",    6.341991,    0.17792659,
  "Conservative", "Traditional",          "Low",     4.904762,    0.18200138,
  "Conservative", "Traditional",          "High",    9.835498,    0.09402925,
  "Liberal",      "Nontraditional",          "Low",     3.570370,    0.13674704,
  "Liberal",      "Nontraditional",          "High",    9.462963,    0.10169496,
  "Liberal",      "Traditional",          "Low",     2.714815,    0.12955408,
  "Liberal",      "Traditional",          "High",    7.644444,    0.13962178
)

# Optional: reorder Happiness levels for plotting
famval_3w_df$Happiness <- factor(famval_3w_df$Happiness, levels = c("Low", "High"))
```


###fig11_fam_3
```{r fig11_familyvalue_threeway}
fam_3_way <- ggplot(famval_3w_df, aes(x = Happiness, y = mean, color = society_type, group = interaction(society_type, poli_ori)))+
  geom_point(size = 3) +
  geom_line(size = .75) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), 
                width = 0.2) +
  scale_y_continuous(breaks = 1:11, limits = c(1, 11)) +
  scale_color_manual(values = c("Nontraditional" = "gray50", "Traditional" = "gray80"))+
  theme_classic() +
  labs(
    x = "Happiness",
    y = "Mean Ideal Society Rating",
    color = "Society Type: Family Values"
  ) +
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 15)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    legend.position = "bottom"
  ) +
  facet_wrap(~poli_ori)


# Print the plot
print(fam_3_way)

# Save the plot



# Save the plot

```

```{r christ_3w_data}
christ_3w_df <- tribble(
  ~poli_ori,      ~society_type, ~Happiness,         ~mean,       ~se,
  "Conservative", "Unimportant",          "Low",     2.816479,    0.13939406,
  "Conservative", "Unimportant",          "High",    6.996255,    0.18727943,
  "Conservative", "Important",          "Low",     4.501873,    0.16487943,
  "Conservative", "Important",          "High",    9.591760,    0.10800660,
  "Liberal",      "Unimportant",          "Low",     3.208696,    0.14105724,
  "Liberal",      "Unimportant",          "High",    9.786957,    0.09826295,
  "Liberal",      "Important",          "Low",     2.426087,    0.11263820,
  "Liberal",      "Important",          "High",    7.286957,    0.18399305
)

# Optional: reorder Happiness levels for plotting
christ_3w_df$Happiness <- factor(christ_3w_df$Happiness, levels = c("Low", "High"))
```

##fig17_ch_3
```{r fig17_christian_threeway}
christ_3w_plot <- ggplot(christ_3w_df, 
                         aes(x = Happiness, y = mean, color = society_type, group = interaction(society_type, poli_ori))) +
  geom_point(size = 3) +
  geom_line(size = .75) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
  facet_wrap(~poli_ori) +
  scale_y_continuous(breaks = 1:11, limits = c(1, 11)) +
  scale_color_manual(values = c("Unimportant" = "gray50", "Important" = "gray80")) +
  theme_classic() +
  labs(
    x = "Happiness",
    y = "Mean Ideal Society Rating",
    color = "Society Type: Importance of Christianity"
  ) +
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 15)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    legend.position = "bottom"
  )

# Print the plot
print(christ_3w_plot)


```
