---
title: "Portfolio 7"
---

#libraries

```{r libraries}
library(tidytuesdayR)
library(tidyverse)
library(ggplot2)
library(forcats)
```


#RollingStone Data
For this portfolio, i am exploring the tidytuesday data on Rolling Stone Album Rankings.


```{r getdata}
rolling_stone <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-05-07/rolling_stone.csv')

```

#convert to long
```{r long}
rolling_stone_long <- rolling_stone %>%
  pivot_longer(
    cols = starts_with("rank_"),
    names_to = "rs_year",
    names_prefix = "rank_",
    values_to = "rank"
  ) %>%
  filter(!is.na(rank)) %>%  
    mutate(rs_year = as.integer(rs_year))

```



#Link to variable info
https://github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-05-07/readme.md

# Exploration of data

## plotting by year

```{r year}
rolling_stone %>%
  ggplot(aes(x = release_year)) +
  geom_histogram() +
  scale_x_continuous(breaks = seq(1950, 2020, by = 5)) +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  labs(title = "Rolling Stone Album Rankings for 2024",
       x = "Release Year",
       y = "Frequency") +
  theme_classic()
```

##diff by year
```{r diff_by_year_plot}
rolling_stone %>%
  ggplot(aes(x = release_year, y = differential)) +
  geom_point() +
  scale_x_continuous(breaks = seq(1950, 2020, by = 5)) +
  scale_y_continuous(breaks = seq(-500, 500, by = 50)) +
  labs(title = "Rolling Stone Album Rankings for 2024",
       x = "Release Year",
       y = "Change in Ranking From 2003 to 2020") +
  theme_classic()
```
##diff by spotify rank
```{r diff_by_spot_plot}
rolling_stone_long %>%
  ggplot(aes(x = spotify_popularity, y = differential)) +
  geom_point() +
  scale_x_continuous() +
  scale_y_continuous(breaks = seq(-500, 500, by = 50)) +
  labs(title = "Rolling Stone Album Rankings for 2024",
       x = "Spotify Popularity",
       y = "Change in Ranking From 2003 to 2020") +
  theme_classic()+
  facet_wrap(~rs_year)
```

## genre
```{r genre}
rolling_stone %>%
  mutate(genre = fct_infreq(genre)) %>%
  ggplot(aes(x = genre)) +
  geom_bar() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(x = "Genre", y = "Number of Albums", title = "Album Counts by Genre")
```

## genre by rs_year
```{r genre_facet}
rolling_stone_long %>%
  mutate(genre = fct_infreq(genre)) %>%
  ggplot(aes(x = genre)) +
  geom_bar() +
  facet_wrap(~rs_year) +  # Facet by year (e.g., 2003, 2012, 2020)
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    strip.background = element_blank(),
    strip.text = element_text(face = "bold")
  ) +
  labs(
    x = "Genre",
    y = "Number of Albums",
    title = "Album Counts by Genre Across Rolling Stone Rankings"
  )
```

## genre by rs_year table
```{r granre_table}
genre_table <- rolling_stone_long %>%
  count(rs_year, genre) %>%
  pivot_wider(names_from = rs_year, values_from = n, values_fill = 0)
print(genre_table)
```

```{r genre_table_mutate}
genre_table <- genre_table %>%
  mutate(diff_2003_2020 = `2020` - `2003`)
```

##violin_year
```{r violine_by_year}
ggplot(rolling_stone_long, aes(x = as.factor(rs_year), y = release_year)) +
  geom_violin(trim = FALSE, fill = "lightblue", color = "darkblue") +  # Create the violin plot
  theme_classic() +  # Use a clean theme
  labs(
    x = "Rolling Stone Rank Year",  # Label for x-axis
    y = "Album Release Year",       # Label for y-axis
    title = "Distribution of Release Year"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for readability
  )
```

##spotify popularity
```{r spotify}
rolling_stone_long %>%
  ggplot(aes(x = rank, y = spotify_popularity)) +
  geom_jitter(width = 0.5, height = 0.5, alpha = 0.5) +
  scale_x_reverse() +
  theme_classic() +
  labs(
    title = "Spotify Popularity by Rolling Stone Album Rank",
    x = "Album Rank (Lower is Better)",
    y = "Spotify Popularity"
  ) +
  facet_wrap(~rs_year)

```

## peak billboard potision
```{r peakbillboardposition}
rolling_stone_long %>%
  ggplot(aes(x = peak_billboard_position, y = spotify_popularity)) +
  geom_jitter(width = 0.5, height = 0.5, alpha = 0.5) +
  scale_x_reverse() +
  theme_classic() +
  labs(
    title = "Spotify Popularity by Rolling Stone Album Rank",
    x = "Peak Billboard Position",
    y = "Spotify Popularity"
  ) +
  facet_wrap(~rs_year)
```

##artist member count
```{r artist_member_count}
ggplot(rolling_stone_long, aes(x = as.factor(rs_year), y = artist_member_count)) +
  geom_violin(trim = FALSE, fill = "lightblue", color = "darkblue") +  # Create the violin plot
  theme_classic() +  # Use a clean theme
  labs(
    x = "Rolling Stone Rank Year",  # Label for x-axis
    y = "Artist Member Count",       # Label for y-axis
    title = "Number of Band Members"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for readability
  )
```
##artist member count X rank
```{r artistnumberandrank}
ggplot(rolling_stone_long, aes(x = rank, y = artist_member_count)) +
  geom_jitter(width = 0.5, height = 0.5, alpha = 0.5) +
  scale_x_reverse() +
  theme_classic() +
  labs(
    title = "Artist Member Count by Rolling Stone Album Rank",
    x = "Album Rank (Lower is Better)",
    y = "Artist Member Count"
  )+
  facet_wrap(~rs_year)
```


```{r lm}
library(dplyr)
library(broom)

# Group by year and run separate models
models_by_year <- rolling_stone_long %>%
  group_by(rs_year) %>%
  nest() %>%
  mutate(model = map(data, ~ lm(rank ~ artist_gender + spotify_popularity + weeks_on_billboard +
                                artist_member_count + release_year, data = .x)),
         results = map(model, tidy))

# View all results
models_by_year %>% select(rs_year, results) %>% unnest(results)
```

#hypothesis
*Might analyze something about how from 2003, 2012, and then 2020 there are shifts in which albums are included??? possible changes in release_year or genre or spotify popularity. Hypothesis: there are more changes (i.e, genre, release year, ranking shift, etc.) between 2012 and 2020 compared to 2003 and 2012 and these shifts are driven by the medium that people are using to listen to music*

I will explore this in my next portfolio piece
